---
title: "Differential expression"
output: html_document
---

```{r, include=FALSE}
library(org.Hs.eg.db)
library(DESeq2)
library(pheatmap)
library(tidyverse)
library(DT)
library(yaml)
library("ggplot2")
library("ggrepel") 
dds <- readRDS("../DEresults.dir/dds.rds")
meta_data <- read.csv("../design.tsv", sep="\t")
source("functions.R")
config <- read_yaml("../pipeline.yml")
```


The following describes the analysis of the data using DEseq2. I have seperated the analysis into the four different contrasts and have plotted summary statistics and MA plots for each.  


## Summary of the data 

```{r}
dds$group <- relevel(dds$group, ref=config$deseq2$control)
res <- results(dds)
summary(res, alpha=config$deseq2$fdr)
```


## MA plots 

In DESeq2, the function plotMA shows the log2 fold changes attributable to a given variable over the mean of normalized counts for all the samples in the DESeqDataSet. Points will be colored red if the adjusted p value is less than 0.01. Points which fall out of the window are plotted as open triangles pointing either up or down.

```{r}
DESeq2::plotMA(res)
```

# Volcano plot

```{r}

res_df <- as.data.frame(res)

res_df <- res_df %>% 
    dplyr::select(log2FoldChange, baseMean, padj) %>% 
    rownames_to_column("GENE") %>% 
  na.omit()
  

mutateddf <- mutate(res_df, sig=ifelse(res_df$padj<0.01, "padj<0.01", "Not Sig")) #Will have different colors depending on significance
input <- mutateddf %>% 
  arrange(mutateddf$padj)

symbol_data <- head(input, 10)

#convert the rownames to a column
volc = ggplot(input, aes(log2FoldChange, -log10(padj))) + #volcanoplot with log2Foldchange versus pvalue
    geom_point(aes(col=sig)) + #add points colored by significance
geom_point(data=symbol_data, aes(log2FoldChange, -log10(padj)), colour="red") +
      ggtitle("MUG KDOBA67") #e.g. 'Volcanoplot DESeq2'

scale_colour_Publication <- function(...){
  library(scales)
  discrete_scale("colour","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)
  
}
volc+geom_text_repel(data=symbol_data, aes(label=`GENE`)) + scale_colour_Publication() + theme_bw()#adding text for the genes

```


## Results table

The folowing results tables show the significant genes. Filtering has been performed with a log2 fold change +/- 2.

```{r , message=FALSE, echo=FALSE}
dir.create("results", showWarnings = FALSE)
rownames(res_df) <- NULL
res_df <- res_df %>% 
  column_to_rownames("GENE")
dt <- filter_genes(as.data.frame(res_df), name="output")

datatable(dt$sig)
```

# Download data

### Full results
<a href="results/output_res.csv" download="results/output_res.csv"><button class="btn"><i class="fa fa-download"></i> Download</button></a>

### Significant results
<a href="results/output_sig.csv" download="results/output_sig.csv"><button class="btn"><i class="fa fa-download"></i> Download</button></a>
